---
title: "ESM 206 HW 3 Task 4"
author: "Sarah Hamilton"
date: "10/24/2021"
output: html_document
---

# Task 4: Naples Reef Purple Urchins

This code contains an exploratory data analysis of purple urchins. It compares
purple urchin size between the kelp removal sites and the no kelp removal sections
of the Naples Reef in 2010.

Data Source: Data Source: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

### Set up

Read in the necessary packages. Hide messages and warnings. 

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)

```

### Read in the data

Read in the data using read_csv() and here(). Clean the column names.

```{r}
urchins_raw <- read_csv(here("data","urchins.csv"))

# Convert the column names to lower snake case using clean_names()

urchins <- clean_names(urchins_raw)
```


### Wrangle the data

Organize the data and make exploratory histograms and QQ plots to analyze the 
size distributions of the annual and control groups of purple urchins. 

```{r}
urchin_subset <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% #filter for purple urchins
  filter(site == "NAPL") %>% #filter for NAPL site
  mutate(date_new = mdy(date)) %>% #make a new data column of class date
  mutate(year_new = year(date_new)) %>% #make a new column for year
  filter(year_new == 2010)
  
```

```{r}
#Make an exploratory histogram

ggplot(data = urchin_subset, aes(x = size)) + #make a histogram
  geom_histogram(aes(fill=treatment)) + #color it by treatment group
  facet_wrap(~treatment) + #separate by treatment group
  labs(x = "Size (cm)", y = "Number of Urchins", 
      title = "Purple Urchin Size Distribution by Control Group") + #add labels
  theme_bw() #add theme
```


```{r}
#Make an exploratory QQ plot 

ggplot(data = urchin_subset) + #make a qqplot 
  geom_qq(aes(sample = size, color = treatment)) + #compare size and color by treatment
  theme_bw() + #add a theme
  facet_wrap(~treatment) + #facet by treatment
  labs(x = "Normal Distribution", y = "Urchin Size", fill = "Treatment Group",
      title = "Purple Urchin Size Distribution by Control Group") 

```

### Analysis of graphs

The annual and control groups for purple urchins seem to have relatively similar
size distributions. They both appear to center around 3-5 cm and both appear
to have ranges of about 2-7 cm. Additionally, they both appear to be relatively
normal, as both QQ plots appear to be mostly linear. 

### Create a summary statistics table

```{r}
#create summary table by grouping by treatment group and summarizing some statistics
summary_table <- urchin_subset %>% 
  group_by(treatment) %>% #group by treatment
  summarize(mean_size = mean(size), #calculate mean
            med_size = median(size), #calculate median
            sd_size = sd(size), #calculate standard deviation
            sample_size = n()) #calculate sample size

#format summary table nicely using kable
final_summary_table <- summary_table %>%
  kable(caption = "Table 1: Summary Statistics for Data on Purple Urchins, separated by Treatment Group", #add caption
        digits = c(2,2,2,2), #reduce table to contain 2 decimal places
        col.names = c("Treatment Group", #add column names
                      "Mean Size (cm)",
                      "Median Size (cm)",
                      "Standard Deviation Size (cm)",
                      "Sample Size")) %>%
  kable_styling(full_width = FALSE)

final_summary_table #output the final summary table
```



