---
title: "ESM 206 HW3 Task 3"
author: "Sarah Hamilton"
date: "10/23/2021"
output: html_document
---

# Santa Barbara Urchins

This code analyzes sea urchin data by comparing the sizes of red and purple
urchins and by visualizing purple urchin populations over time at various sites.

Data Source: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

### Set up 

Hide warnings and messages, show code.
Read in necessary packages. Hide messages and warnings. 

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)

```

### Read in the data

Read in the data using read_csv() and here(). Clean the column names.

```{r}
urchins_raw <- read_csv(here("data","urchins.csv"))

# Convert the column names to lower snake case using clean_names()

urchins <- clean_names(urchins_raw)
```

# Section 1: Size comparisons

Size comparisons for purple and red urchins.

### Data Visualization

Compare the size distribution of purple and red urchins using a histogram. 
Use 10-15 bins. 
Add a figure caption. 

```{r}
ggplot(data = urchins, aes(x = size)) + #plot based on size
  geom_histogram(bins = 12, aes(fill = common_name)) + #create a histogram with color based on common_name
  theme_bw() + #add a bw theme
  facet_wrap(~common_name) + #sort into two graphs by common_name()
  labs(x = "Size (cm)", 
       y = "Count", title = "Size Distribution of Purple and Red Urchins", 
       fill="Legend",
       caption = "Figure 1. Size distributions of purple and red urchins. The size distribution for purple urchins \n is shown on the left and the size distribution for red urchins is shown on the right.") + #add labels
  scale_fill_manual(values = c("purple","red")) #change colors to coordinate with type of urchin
  
```

### Descriptive Statistics Table

Create a finalized summary table that contains the mean, standard deviation,
and sample size for purple and red urchins.

```{r}
sum_table <- urchins %>%
  group_by(common_name) %>% #group by common name
  summarize(mean_size = mean(size), #calculate mean
            median_size = median(size), #calculate median
            std_dev_size = sd(size), #calculate standard deviation
            sample_size = n() #calculate sample size
            )

final_table <- sum_table %>%
  kable(caption = "Table 1: Summary Statistics for Data on Sea Urchins.", #add table caption
        digits = c(2,2,2,2), #display two decimals in the table
        col.names = c("Color of Urchin", #format the column names
                      "Mean Size (cm)", 
                      "Median Size (cm)", 
                      "Standard Deviation Size (cm)",
                      "Sample Size")) %>%
  kable_styling(full_width = FALSE)

final_table
```


### Analysis of Results

Brief description of the differences in the summary statistics for the 2 different types of urchins:

Red urchins are about 2.3 cm (or about 50%) bigger than the purple urchins on average, and the red
urchins have a standard deviation that is about 0.7 cm (or about 60%) greater than that 
of the purple urchins. These statistics along with the shape of the two graphs indicate 
that purple sea urchins tend to be both smaller and slightly more uniform in size than 
red sea urchins. 

# Section 2: Purple Urchin Counts by Site and Year

Create a subset of urchins data. 

```{r}
purple_urchins <- urchins %>%
  filter(common_name == "Purple Urchin") %>% #filter for purple urchins
  mutate(date_new = mdy(date)) %>% #make a new data column of class date
  mutate(year_new = year(date_new)) %>% #make a year column
  mutate(site_full = case_when( #make a new column with full names 
    site == "NAPL"~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))

```

Store a table containing total number of purple urchins observed by site and year.

```{r}
purple_site_year <- purple_urchins %>%
  group_by(site_full, year_new) %>% #gropu the data by site and year
  summarize(count_urchins = n()) #total the number of urchins at each site in each year

purple_site_year
```

Make a graph of counts of purple urchins by site over time.

```{r}
ggplot(data = purple_site_year, aes(x = year_new, y = count_urchins)) + #plot the counts over time for each site
  geom_col(aes(fill = site_full)) + #column graph 
  facet_wrap(~site_full) + #facet wrap to split up the graphs by site
  theme_minimal() +
  labs(x = "Year", y = "Number of Urchins", title = "Purple Urchin Population Over Time",
       fill = "Site",
       caption = "Figure 2. Purple Sea Urchin Populations Over Time at 5 Sites in Santa Barbara")
```

Brief analysis of graph:

This graph seems to indicate that there is a decrease in the purple sea urchin
population along the Santa Barbara coast, because at each site the count
of sea urchins seems to have decreased significantly since the early 2010s.
However, to determine whether the counts reflect actual changes in population,
we would need to know whether the sea urchins were simply moving away from these
sites to other sites along the coast or whether they are dying as a result of
kelp removal. The only way to know this would be to survey many other sites to
have more data or to know the whole population, both of which are very hard or 
impossible to do. Therefore, while this data indicates lower urchin populations
at these specific sites, we cannot say definitively what is happening to these
urchins.






